Supporting Information

### 2018-04-02 SupportingInformation.Rmd
minor changes
edited some more in the aphid sections how that shya has fixed some plots and added accompanyng text. Can next start from plankton.

## Abundance data for *Ceratium furca* in UK seas \label{Pns_data}
   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
   * We threw out all locations that had fewer than 45 years data
   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
   * *Ceratium furca* is a dicoflagellate commonly found in UK seas
   * note, when we edit this, we cannot call it the North Seas since data come from around the UK not only the North Sea, unless Shya threw out all grid cells that were outside the North Sea

@@

<!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
\begin{figure}[!h]
\begin{center}
\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (E)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
\caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance $(D)$ between two sampling locations; $\textbf{(A)}$ Spearman correlation vs. $D$, $\textbf{(B)}$ Kendall correlation vs. $D$, $\textbf{(C)}$ $\cor_l-\cor_u$ vs. $D$, $\textbf{(D)}$ $\Ps_l-\Ps_u$ vs. $D$, $\textbf{(E)}$ $\Dsq_u-\Dsq_l$ vs. $D$ plotted for all (10 $\times$ 10 locations) taken pairwise. \label{fig_aphid_count}}
\end{center}
\end{figure}

We computed Spearman and Kendall correlation along with our three non-parametric statistics ($\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$) as described in Sec. \ref{NPA} for  green spruce aphid abundance data taken at all $10$ $\times$ $10$ location pairs (code given in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R)). We also constructed a $10$ $\times$ $10$ matrix using `gcdist` from `ncf` package where each element in that matrix represented the geographical distance between any two selected locations. 
Figure \ref{fig_aphid_count} presented plots for **(A)** Spearman, **(B)** Kendall, **(C)** $\cor_l-\cor_u$, **(D)** $\Ps_l-\Ps_u$ and **(E)** $\Dsq_u-\Dsq_l$ statistics against the geographical distance for all selected sampling location pairs. With increasing the distance *D*, the Spearman and Kendall correlation (hence the synchrony among time series data) between the location pairs decreased as expected. But one cannot conclude the tail dependence of that synchrony from Figure \ref{fig_aphid_count} **(A)** or **(B)**, whereas **(C - E)** clearly showed the lower tail dependence with most of the data points lied in the positive quadrant as described before in Sec. \ref{NPA}.

@@

@@

<!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
\begin{figure}[!h]
\begin{center}
\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (E)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
\caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations; $\textbf{(A)}$ Spearman correlation vs. $D$, $\textbf{(B)}$ Kendall correlation vs. $D$, $\textbf{(C)}$ $\cor_l-\cor_u$ vs. $D$, $\textbf{(D)}$ $\Ps_l-\Ps_u$ vs. $D$, $\textbf{(E)}$ $\Dsq_u-\Dsq_l$ vs. $D$ plotted for all (10 $\times$ 10 locations) taken pairwise. \label{fig_aphid_ff}}
\end{center}
\end{figure}

Figure \ref{fig_aphid_ff} represented plots for **(A)** Spearman correlation, **(B)** Kendall correlation, **(C)** $\cor_l-\cor_u$, **(D)** $\Ps_l-\Ps_u$ and **(E)** $\Dsq_u-\Dsq_l$ statistics, computed on leaf-curling plum aphid's first flight date data, against the geographical distance for all selected sampling location pairs. Likewise in Figure \ref{fig_aphid_count}, with increasing the distance *D*, the Spearman and Kendall correlation (hence the synchrony among time series data) between the location pairs decreased as expected. But here in contrast, Figure \ref{fig_aphid_ff}  **(C - E)** clearly showed the upper tail dependence with most of the data points lied in the negative quadrant as described before in Sec. \ref{NPA}.

@@

\newpage
## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

@@

\newpage
# References:

>>>>

## Abundance data for *Ceratium furca* in UK seas \label{Pns_data}
   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
   * We threw out all locations that had fewer than 45 years data
   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
   * *Ceratium furca* is a dicoflagellate commonly found in UK seas
   * note, when we edit this, we cannot call it the North Seas since data come also from other seas around the UK

@@

@@

Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance
between the locations (Figure \ref{fig_aphid_count} A,B) showed the typical decline of correlation with 
distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of  
locations shows these quantities also decline with distance (Figure \ref{fig_aphid_count} C-H).
<!--DAN add reference and revisit the text here once Shya add these panels-->
Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Figure \ref{fig_aphid_count} I-K)
did not show a decline with distance, but did illustrate visually the statistical
result of Table \ref{tab_count_resamp} that averages of these quantities across all pairs of sampling locations
were significantly greater than zero, demonstrating asymmetry of tail dependence.
Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in 
the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.

<!--Shya, can you please add panels for corl, coru, Pl, Pu, D2l, D2u vs. distance, following the Spearman and Kendall
panels? I think it will be the first time anyone has shown these, so we should show the decline with distance
holds for these as well. That means you'll have 6 new panels, for a total of 11 panels. So it'll be a big figure, 
possibly taking up a whole page, or most of one. Similar changes, please,
to the analogous figs for the aphid ff and ceratium furca data.-->
\begin{figure}[!h]
\begin{center}
\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (E)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
\caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance ($D$) between two sampling locations. (A) Spearman correlation vs. $D$; (B) Kendall correlation vs. $D$; (C) $\cor_l$ vs. $D$; (D) $\cor_u$ vs. $D$; (E) $\Ps_l$ vs. $D$; (F) $\Ps_u$ vs. $D$; (G) $\Dsq_u$ vs. $D$; 
(H) $\Dsq_l$ vs. $D$; 
(I) $\cor_l-\cor_u$ vs. $D$, (J) $\Ps_l-\Ps_u$ vs. $D$, (K) $\Dsq_u-\Dsq_l$ vs. $D$ plotted for all 
pairs of locations. \label{fig_aphid_count}}
\end{center}
\end{figure}

@@

@@

Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Figure \ref{fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance. Plotting $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$, and $\Dsq_l$ against distances between pairs of locations shows these quantities also decline with distance (Figure \ref{fig_aphid_ff} C-H).
<!--DAN: make sure this is true when you see the plots again-->
Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.

<!--Shya, pls add 6 new panels as described in my comment near the analogous figure for aphid count-->
\begin{figure}[!h]
\begin{center}
\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}\\
\textbf{\;\;\;\;\;\;\;\ (E)} \\
\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
\caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations. (A) Spearman correlation vs. $D$, (B) Kendall correlation vs. $D$, (C) $\cor_l$ vs. $D$, (D) $\cor_u$ vs. $D$; (E) $\Ps_l$ vs. $D$; (F) $\Ps_u$ vs. $D$; (G) $\Dsq_u$ vs. $D$; (H) $\Dsq_l$ vs. $D$; (I) $\cor_l-\cor_u$ vs. $D$; (J) $\Ps_l-\Ps_u$ vs. $D$; (K) $\Dsq_u-\Dsq_l$ vs. $D$ plotted for all pairs of locations. \label{fig_aphid_ff}}
\end{center}
\end{figure}

<!--DAN: effectively stopped here 2018 04 01 10pm-->

## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

@@

# References:
